---
title: "lab_03"
author: "Derek Willis"
date: "2023-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse

# Load janitor
```

## TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans AND, using janitor, change the column names so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
library(readr)
State_of_Maryland_Payments_Data_FY2008_to_FY2024 <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv")
View(State_of_Maryland_Payments_Data_FY2008_to_FY2024)
# Load in the Maryland grants and loan data
library(readr)
State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022 <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
View(State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022)
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed in the payments data. Look at & search the records in the dataframe to start your search, starting with a broad search and then narrowing it down. Then write code to find rows where the vendor name contains some version of the county's school system. 

HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find? 

A1: I found 6 separate Grantee's that associated with the Prince George's County school system being paid. 

```{r}
library(tidyverse)

# find all of the variations of how the school system might be listed in the payments data
specific_name <- c("PRINCE GEORGES CHILD RESOURCE CENTER INC", "Prince Georges Child Resource Center Inc", "IMAGINE PRINCE GEORGES COUNTY LLC", "PRINCE GEORGES COMMUNITY COLLEGE", "PRINCE GEORGES CO DEPT OF FAMILY SERVICES", "PRINCE GEORGES COUNTY DEPT OF FAMILY SERVICES")

State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022 |>
  group_by("PRINCE GEORGES CHILD RESOURCE CENTER INC", "Prince Georges Child Resource Center Inc", "IMAGINE PRINCE GEORGES COUNTY LLC", "PRINCE GEORGES COMMUNITY COLLEGE", "PRINCE GEORGES CO DEPT OF FAMILY SERVICES", "PRINCE GEORGES COUNTY DEPT OF FAMILY SERVICES") |>
  arrange(desc(State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022))
  


  
 


```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2: 2020 had the largest total amount. One part of me is surprised that the school system was able to receive the most amount of grants and loans during a time where the economy was at an all time low due to the inability to work... But another part of me is not so shocked considering it would be a time like then where people will come together to keep education flourishing...

```{r}
# Group the data by fiscal year and calculate the total payment amount for each year
library(tidyverse) 

State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022 |>
  arrange(desc(`Fiscal Year`)) |>
  count(Amount, "PRINCE GEORGES CHILD RESOURCE CENTER INC", "Prince Georges Child Resource Center Inc", "IMAGINE PRINCE GEORGES COUNTY LLC", "PRINCE GEORGES COMMUNITY COLLEGE", "PRINCE GEORGES CO DEPT OF FAMILY SERVICES", "PRINCE GEORGES COUNTY DEPT OF FAMILY SERVICES")
#
library(lubridate)
State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022 |> 
  group_by(`Fiscal Year`, Category, "PRINCE GEORGES CHILD RESOURCE CENTER INC", "Prince Georges Child Resource Center Inc", "IMAGINE PRINCE GEORGES COUNTY LLC", "PRINCE GEORGES COMMUNITY COLLEGE", "PRINCE GEORGES CO DEPT OF FAMILY SERVICES", "PRINCE GEORGES COUNTY DEPT OF FAMILY SERVICES") |> 
  summarize(Amount = n()) |> 
  arrange(`Fiscal Year`)


```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. 

Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe in words how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. I first had to make an arrange function to show me solely each grant or loan provided by each year. Then I cleaned it up to where I only saw payments made towards 'Capital Projects- University of Maryland'. Mitchell Dundon payments, Cianbro Corporation payments, and Crown Castle Fiber LLC appear multiple times amongst various Agency names, those should be grouped together possibly? Maybe I would want to arrange it so where its in the descending order of the last name of the Vendor names?

```{r}
library(tidyverse)
library(lubridate)

State_of_Maryland_Payments_Data_FY2008_to_FY2024 |>
  arrange(desc('Amount')) |>
  group_by("CAPITAL PROJECTS- UNIVERSITY OF MARYLAND")
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which column(s) contain the information you want, your code should find just the grants for stem cell research in a single filter. Then calculate the total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: To summarize how stem cell research funding has changed over time, by each year, we notice more separate donations by different entities. I feel like it goes to show that stem cell research became something that was worth giving more donations as more research showed provided there was more to discover. 

```{r}
library(tidyverse)
library(lubridate)

State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022 |>
  arrange(desc('Amount')) |>
  group_by("Stem Cell Research Grant, STEM CELL RESEARCH GRANT")

```

Q5: Using the Maryland grants and loan data, find records where the zip code is the one where the University of Maryland-College Park is located in and produce a list of grantees from that zip code, along with the total amount they received. Write a paragraph explaining how many grantees are there, what you would need to do to clean the results so that you had one record per grantee and which grantees are most interesting to you?

A5: It starts with identifying that the University of Maryland College Park's zip code is 20742. With that I arranged the data by the zip code. With that I was able to see that there were 9 different Grantees. I think you would have to find a way to combine all of the like Grantee donations together. But, is there a way to do so, so it separated by the year? That way it could make for more interesting observations pertaining to the trend of the total amount by the year. 

```{r}
library(tidyverse)
library(lubridate)

State_of_Maryland_Grant_and_Loan_Data_FY2009_to_FY2022 |>
  arrange(desc('Zip Code')) |>
  group_by("University of Maryland")

```
