---
title: "lab_11"
author: "derek willis"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries for working with data, plus rvest.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
```


Let's get to scraping.

## Questions

**Q1**. Scrape the listing of available Maryland state grants at https://grants.maryland.gov/Pages/StateGrants.aspx into a dataframe. You should have three columns, one of which is a date, so make sure the date column has a date datatype. Then, write code to count the number of grants opportunities offered by each organization listed in your dataframe, showing the organization with the most grant opportunities first. Which state agency has the most?

**A1** Maryland Energy Administration has the most grant opportunities

```{r}
state_grants_url <- "https://grants.maryland.gov/Pages/StateGrants.aspx"
```

```{r}
state_grants_url_results <- state_grants_url |>
  read_html()

state_grants_url_results
```

```{r}
state_grants_url_results <- state_grants_url |>
  read_html() |>
  html_table()

# show the dataframe

state_grants_url_results
```

```{r}
state_grants_url_results <- state_grants_url |>
  read_html() |>
  html_table() |>
  bind_rows() |>
  clean_names() |>
  mutate(due_date =ymd(due_date)) |>
  group_by(organization) |>
  summarise(count=n()) |>
              arrange(desc(count))

state_grants_url_results
  
  
```

**Q2** Next, let's scrape the list of press releases from Maryland's Office of the Public Defender, https://www.opd.state.md.us/press-releases. This isn't a table, so you'll need to use `html_elements()` and your browser's inspector and do some clean up on the results. The result should be a dataframe with two columns that contain the date and title, and the date column should have a date datatype. The challenge here is figuring out how to isolate the releases.

When you finish scraping into a dataframe, write code to find the press releases that have the word "police" in the title. How many are there and when was the most recent one?

```{r}
press_releases_url <- "https://www.opd.state.md.us/press-releases"
```

```{r}
press_releases_url_results <- press_releases_url |>
  read_html()

press_releases_url_results
```

```{r}
press_releases_url_results <- press_releases_url |>
 read_html() |>
  html_elements('span.wixui-rich-text__text') |>
  html_text() |>
  as_tibble() |>
  clean_names() |>
  separate(value, c("date", "title"), sep=":") |>
  mutate(date = as.Date(date, format = "%m/%d/%Y")) |>
  filter(!is.na(title),
         str_detect(title, "Police")) |>
  group_by(title, date) |>
  summarise(count = n()) |>
  arrange(date)
  

press_releases_url_results

```

**A2** There are 9 with Police in the title... The most recent was June 21, 2021.

**Q3** Sen. Ben Cardin, D-Maryland, has posted hundreds of press releases at https://www.cardin.senate.gov/?post_type=press-releases. It would be great to have all of them in a dataframe that has the following columns: date, title and url.

To do this, you will need to scrape the page's html and save that to a variable, and _then_ extract the dates, titles and urls into _separate_ dataframes using html_elements(). And remember how we turn a list into a dataframe. The function `html_text()` pulls out the contents of a tag, but for urls we want the HTML attribute. Rvest gives you a way to extract the URL from a link; google to find out what it is.

At the end, you'll have three dataframes that you want to combine into a single dataframe. When we want to combine the rows of identical dataframes, we used `bind_rows()`. If you were combining columns instead of rows, there's a similar function. Find out what it is and use it to put all of the dataframes together into a single one.

When you're done, rename the columns so they make sense, then make sure the date column is an actual date.

Finally, tell me what questions you could ask of this data. Be creative.

```{r}
releases_url <- "https://www.cardin.senate.gov/?post_type=press-releases"
```

```{r}
releases_url_results <- releases_url |>
  read_html()

releases_url_results
```

```{r}
date <- releases_url_results |>
  html_elements('h5.customBlog_item__date') |>
  html_text() |> 
  as_tibble() |>
  rename(date = value)

date$date <- mdy(date$date)

date
```


```{r}
title <- releases_url_results |>
html_elements('h3.customBlog_item__title') |>
  html_text() |> 
  as_tibble() |>
  rename(title = value)
  
title$title <- gsub("^\\n", "", title$title)

title
```


```{r}
url  <- releases_url_results |>
  html_elements('a.customBlog_item__readMore') |>
  html_attr('href') |> 
  as_tibble() |>
  rename(url = value)
url
```

**A3** 
Regarding the statement about the release of Hostages of Gaza.. I would ask or try to research about the reasoning for the release of the hostages... And what are the next steps that need to be done to urge the President about how to find them??

On the announcement regarding over $2 million to strengthen infrastructure resiliency at the City of Havre de Grace... Maybe you could ask to see a report, or just a more structured breakdown of where exactly the money is going to.. Specific buildings, maybe they are hospitals, quick homes for families with damaged homes due to tragic events... And from there it would allow for concerned readers to know more information?


```{r}
combined <- cbind(date, title, url)

combined
```
