---
title: "lab_05"
author: "derek willis"
date: "2023-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

- Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
install.packages("tidyverse")
library(tidyverse)
library(lubridate)
library(dplyr)
library(janitor)
```

## Get Our PDF

We'll be working with the [911 overdose calls from Baltimore County](https://drive.google.com/file/d/1qkYuojGF_6WKFr5aNQxmewDzcKyOiJFr/view?usp=share_link). You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to extract the tables within it, export that to a CSV file, load it into RStudio and ask some questions.

## Extract Data from PDF Using Tabula

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds or longer.

This PDF has a single table spread over multiple pages to extract. We're going to make a single dataframe from this table, exporting it to a CSV file that you will load into R. In Tabula, highlight the table and click the "Preview & Export Extracted Data" button. You may want to play with including or excluding the column headers - YOU SHOULD HAVE FIVE COLUMNS OF DATA.

Save the CSV (it should be called `tabula-Baltimore County; Carey, Samantha log OD.csv` by default) to your lab_05/data folder.

From there, you will need to read in the data, and add or fix headers if necessary. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)).

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers should not contain spaces and they should have meaningful names, not "x1" or something similar. How you do that is up to you, but you can use select() with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns. Importantly, you'll need to ensure that any columns containing a date actually have a date datatype. Our friend `lubridate` can help with this.

```{r}

```


```{r}
# Load your data into a data frame (replace "data" with your data frame name)
data <- read.csv("tabula-Baltimore County; Carey, Samantha log OD.csv")

# Remove spaces from column names
colnames(data) <- gsub(" ", "", colnames(data))

# Load your data into a data frame (replace "data" with your data frame name)
data <- read.csv("tabula-Baltimore County; Carey, Samantha log OD.csv")

# Assign new column names without spaces
new_colnames <- c("DATE", "CALLTIME/CASENUMBER", "ODTYPE","LOCATION")
colnames(data) <- new_colnames

head(data)

```

## Answer questions

Q1. Write code to generate the number of calls that occurred on each date. Which date in 2022 had the most overdose calls, and how many? Look at the total number of rows in your result and explore the range of dates - based on your result, do you believe there are any days with no overdose calls at all? Explain why or why not.

A1. 07/14/22 and 10/04/22 had the most overdose calls, with 23 each.. I think there are some days with zero... Solely because any date that didn't have a call, should not and would not be present in this set of data.

```{r}
# Create a frequency table of unique dates and their counts
date_counts <- table(data$DATE)

#
date_counts_data <- as.data.frame(date_counts)

#
colnames(date_counts_data) <- c("DATE", "Count")

#
print(date_counts_data)

#
date_counts_data <- date_counts_data[order(-date_counts_data$Count), ]

#
print(date_counts_data)

```

Q2. You want to understand if there's a pattern in the day of the week that overdose calls are made. Add a column to your dataframe that displays what day of the week each date represents. You should search for how to do that using lubridate. Then write code to calculate the number of calls for each day of the week, and add a column to that result that calculates the percentage of all calls that occurred on each day of the week (so you want a dataframe with the day of the week, total number of calls and the percentage of calls on that day out of the total number of all calls). Describe your findings to me.

A2. 

```{r}
library(lubridate)
library(dplyr)
```

```{r}
# Convert the "DATE" column to character and then to Date
date_counts_data$DATE <- as.Date(as.character(date_counts_data$DATE), format = "%m/%d/%y")

# Add the "DayOfWeek" column
date_counts_data$DayOfWeek <- weekdays(date_counts_data$DATE)

# Calculate the number of overdose calls for each day of the week
day_of_week_counts <- date_counts_data %>%
  group_by(DayOfWeek) %>%
  summarize(TotalCalls = n())

# Calculate the percentage of calls for each day of the week relative to the total number of all calls
day_of_week_counts <- day_of_week_counts %>%
  mutate(Percentage = (TotalCalls / sum(TotalCalls)) * 100)

# Display the resulting data frame
print(day_of_week_counts)






```


Q3. Now let's look at locations. Which ones have the most calls? How would you describe them (feel free to search for more information on them)? Is there anything about the structure of the original data that might make you less confident in the counts by location or date?

A3.  I am seeing Liberty Road having the most calls... I would say that this specific road located in Baltimore had a serious issue with overdoses due to the amount of calls in the data. It is fair to assume that illegal and highly dangerous drugs are being distributed in a high volume to where there are multiple counts of ODs right on that street. It could be my fault but I am getting what it seems to be zip codes or area codes within my location? which makes me much less confident.

```{r}
library(dplyr)

library(dplyr)

# Assuming your data frame is named "DATA" and the location data is in the "LOCATION" column
# Group the data by "LOCATION" and count the occurrences
location_counts_date_counts_data <- data %>%
  group_by(LOCATION) %>%
  summarize(Count = n())

# Sort the result from lowest to highest counts
location_counts_date_counts_data <- location_counts_date_counts_data %>% arrange(Count)

# Display the result
print(location_counts_date_counts_data)



```

Q4. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A4. Why is Liberty Road getting so many calls pertaining to ODs? How close are they to a hospital, or specific area so the access to the cause of these ODs are at such a high volume?
